\vspace{24pt} Comparison of Different Data Layouts for Columnar In-Memory Databases on the Basis of an Example Application

### 2018-08-31 paper/paper.tex
add figure reference
add benchmarks: coordinates and time as multilpe columns

As the Frame format is geared to such time frames it should be much faster to get all points within a frame. With 14 milliseconds it is faster than the Points format, but only by a small amount. The reason for that is the delta encoding for all but the first coordinate within a frame. So for each found entry the coordinate needs to be calculated. After this is done, rounding the coordinates according to the granulatrity and grouping to get a weight is the same as for the Points format.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{img/bench_single_frame.png}
  \caption{Time in ms required for retrieving one single time frame}
  \label{fig:bench_whole_trajectory}
\end{figure}

@@

>>>>

As the Frame format is geared to such time frames it should be much faster to get all points within a frame. Figure \ref{fig:bench_whole_trajectory} shows that with 14 milliseconds it is faster than the Points format, but only by a small amount. The reason for that is the delta encoding for all but the first coordinate within a frame. So for each found entry the coordinate needs to be calculated. After this is done, rounding the coordinates according to the granulatrity and grouping to get a weight is the same as for the Points format.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{img/bench_single_frame.png}
  \caption{Time in ms required for retrieving one single time frame}
  \label{fig:bench_whole_trajectory}
\end{figure}

@@

\subsubsection{Storing floats as two integers}
Floating point operations tend to be slower than integer operations. We split up each float column for latitude and longitude into two integer columns with the intention to speed up calculations. For the test we filtered all entries to be within a given bounding rectangle. In fact, this aproach more than doubled the processing time of the query for multiple integer columns compared to the single float ones as depicted in Figure \ref{fig:bench_coordinates_as_int}.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{img/bench_coordinates_as_int.png}
  \caption{Time in ms filtering by coordinates}
  \label{fig:bench_coordinates_as_int}
\end{figure}

\subsubsection{Storing time in multiple columns}
One successful optimization was to store time as seconds instead of timestamps. But when filtering by time often the hours, minutes or seconds part of a time data point is important. With that in mind, we stored the timestamps in three colums for the hour, minutes and seconds part with the purpose of better access and faster filtering. However, as illustrated in Figure \ref{fig:bench_time_as_multiple_ints} our measurements showed that filtering by time became three times slower than with just using seconds.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{img/bench_time_as_multiple_ints.png}
  \caption{Time in ms required for filtering by time}
  \label{fig:bench_time_as_multiple_ints}
\end{figure}

\subsubsection{Get taxis in time and range boundary}
